---
title: "Ueberhangmandate 2017"
output: rmarkdown::github_document
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(
  fig.path = "README_figs/"
)
```

This is the repository for the blog post published as "TITLE" on "DATE". It contains data from the Bundeswahlleiter.


## Folder Layout
- the blog_text.Rmd is used to create the blog post
- the scripts folder contains the r scripts to create the data and graphs, which are stored in data and images respectively.
- the dashboard folder contains the dashboard (the data for which is created with the 01_data_cleaning.R in the scripts folder).

## Scripts
- 01_data_cleaning.R
- 01_parliament_size_calculation.R (this is used in '01_data_cleaning.R')
- 02_graphs.R
- 02_themes (this is used in '02_graphs.R')
- 03_app (the script can be found in the dashboard folder)

## Packages needed for the project
```{r , eval = FALSE}
library(dplyr) # data manipulation
library(tidyr) # data manipulation
library(stringr) # data manipulation
library(rgdal) # for reading in shape file
library(lpSolveAPI) # optimization
library(rvest) # web scrapping
library(stringr) # structure replacement Überhang to Ueberhang etc.
library(knitr) # package for tables
library(janitor) # for better variable names
#--------------------
library(grid) # plots
library(ggplot2) #plots
library(ggpubr) # plots next to each other
library(rgdal) # for maps
library(rgeos) # for maps
library(mapproj) # for maps
library(gganimate) # for gif
library(transformr) # for gif
#--------------------
library(shiny) # for the dashboard
library(shinydashboard) # for the dashboard
library(shinyWidgets) # for the dashboard
```


# Licence 
The raw election data is published under Data licence Germany – attribution – version 2.0. 
https://www.govdata.de/dl-de/by-2-0. It can be downloaded from: https://www.bundeswahlleiter.de/dam/jcr/72f186bb-aa56-47d3-b24c-6a46f5de22d0/btw17_kerg.csv.
In this repository it can be found under 'data/raw/raw.csv'. The data for the historic size of the parliament can be found under: https://www.bundeswahlleiter.de/service/glossar/u/ueberhangmandate.html.
All other election data was modified by the author.

The shape files where downloaded from the Berliner Morgenpost (https://interaktiv.morgenpost.de/analyse-bundestagswahl-2017/data/btw17-shapes.zip).






