---
title: "Ueberhangmandate"
output: rmarkdown::github_document
---
  
  ```{r setup, include=FALSE}
knitr::opts_chunk$set(
  fig.path = "README_figs/"
)
```

## Packages

```{r packages, message=FALSE}
# data manipulation
library(tidyr)
library(dplyr)
# optimization
library(lpSolveAPI)
#web scrapping
library(rvest)
#structure replacement Überhang to Ueberhang etc.
library(stringr)
# plots
library(grid)
library(ggplot2)
# packages for maps
library(rgdal)
library(rgeos)
library(mapproj)
# package for table
library(knitr)
```

```{r installs,echo=FALSE, eval=FALSE}
# installs needed in ubuntu 18.04
sudo apt-get install gdal-bin proj-bin libgdal-dev libproj-dev
```


## Theme
```{r theme}
hp_theme <- function(base_size = 13, base_family = "") {
  theme_grey(base_size = base_size, base_family = base_family) %+replace%
    theme(
      
      # Base elements which are not used directly but inherited by others
      line =              element_line(colour = '#DADADA', size = 0.75,
                                       linetype = 1, lineend = "butt"),
      rect =              element_rect(fill = "#F0F0F0", colour = "#F0F0F0",
                                       size = 0.5, linetype = 1),
      text =              element_text(family = base_family, face = "plain",
                                       colour = "#656565", size = base_size,
                                       hjust = 0.5, vjust = 0.5, angle = 0,
                                       lineheight = 0.9, margin = margin(), debug = FALSE),
      plot.margin =       margin(12,10,5,10),
      # Modified inheritance structure of text element
      plot.title =        element_text(size = rel(0.75), family = '' ,
                                       face = 'bold', hjust = 0,
                                       vjust = 2.5, colour = '#3B3B3B'),
      plot.subtitle =     element_text(size = rel(0.4), family = '' ,
                                       face = 'plain', hjust = 0,
                                       vjust = 2.5, colour = '#3B3B3B', margin = margin(0,0,15,0)),
      axis.title.x =      element_blank(),
      axis.title.y =      element_blank(),
      axis.text =         element_text(),
      # Modified inheritance structure of line element
      axis.ticks =        element_line(),
      panel.grid.major =  element_line(),
      panel.grid.minor =  element_blank(),
      
      # Modified inheritance structure of rect element
      plot.background =   element_rect(),
      panel.background =  element_rect(),
      legend.key =        element_rect(colour = '#DADADA'),
      
      # Modifiying legend.position
      legend.position = 'none',
      
      complete = TRUE
    )
}
```

## Historic Size Bundestag Data Set

```{r historic_size_data, error=FALSE, message=FALSE, warning=FALSE}
# read in source
  url <- 'https://www.bundeswahlleiter.de/service/glossar/u/ueberhangmandate.html'
# as the data is presented as two tables on the url we need to read them in separately
# first table [until 2009](before current election reform): no adjustments for Überhangmandate
  first_part = url %>%
    read_html() %>%
    html_node(xpath = '/html/body/div/div/main/table[1]') %>%
    html_table(fill = TRUE)
# clean up: first row contains names needed in the variable names. Then apply
# consistent and more intuitive names and keep only relevant variables 
  first_part[] <- lapply(first_part, as.character)
  names(first_part) <- paste(names(first_part), first_part[1, ], sep = "_")
  first_part <- first_part %>%
                    rename(`Jahr`=`Jahr der Bundestagswahl_Jahr der Bundestagswahl`,
                          `Wahlkreise`=`Zahl der Wahlkreise_Zahl der Wahlkreise`,
                          `Mandate`=`Sitze insgesamt 1_Sitze insgesamt 1`) %>%
                  slice(2:49) %>%
                  select(c(-5)) %>%
                  rename_all(funs(str_replace(., "Überhangmandate", "Ueberhangmandate")))

# second table [2013,2017](after current election reform): adjustments for Überhangmandate
  second_part = url %>%
    read_html() %>%
    html_node(xpath = '/html/body/div/div/main/table[2]') %>%
    html_table(fill = TRUE)
# clean up (see above)
  second_part[] <- lapply(second_part, as.character)
  names(second_part) <- paste(names(second_part), second_part[1, ], sep = "_")
  second_part <- second_part %>%
                  rename(`Jahr`=`Jahr der Bundestagswahl_Jahr der Bundestagswahl`,
                        `Wahlkreise`=`Zahl der Wahlkreise_Zahl der Wahlkreise`,
                        `Mandate`=`Sitze insgesamt_Sitze insgesamt`) %>%
                  rename_all(funs(str_replace(., "Überhänge", "Ueberhangmandate"))) %>%
                  slice(c(2:8)) %>%
                  select(c(1:6))  

# merge data frames and make everything numeric as it's needed for the graph 
  historic_size <- full_join(first_part,second_part)
  historic_size[, 1:5] <- sapply(historic_size[, 1:5], as.numeric)
  historic_size <- historic_size%>% mutate(Ueberhangmandate_Partei=replace(Ueberhangmandate_Partei, Ueberhangmandate_Partei=="–", NA)) 
# drop unnecessary data
  rm(first_part,second_part, url)
```

## Historic Size Bundestag Graph
```{r historic_size_graph}
# as the data frame contains every Ueberhangmandat per party, I only need one observation per year
  historic_size %>%
    distinct(`Jahr`, .keep_all = TRUE) %>%
    select(c(`Jahr`,`Mandate`)) %>%
    ggplot(aes(x=`Jahr`, y=`Mandate`,group=1)) +
    geom_line(aes(group=1), color="#009E73") +
    geom_hline(aes(yintercept=598)) +
    scale_y_continuous(limits = c(300, 800.1),
                       breaks = c(300,400,500,598,700,800), 
                       expand = c(0, 0),
                       labels=c("300" = "300", "400"="400", "500"="500", "598"="598 \n (heutige Normgröße)","700"= "700","800"="800 Sitze")) +
    scale_x_continuous(breaks = sort(c(seq(1960, 2010, by = 10),1949,2017)),
                       limits=c(1949,2017), 
                       labels=c("1949"="1949","1960"="60","1970"="70","1980"="80","1990"="90","2000"="2000","2010"="10", "2017"="17")) +
    labs(title = "Historische Entwicklung der Größe des Bundestages", subtitle="",caption = "Quelle: Bundeswahlleiter") +
    hp_theme() + theme(axis.text= element_text(size=7.5), axis.title.x = element_blank(),plot.title.position = "plot",  axis.title.y = element_blank(), 
                       panel.grid.major.x = element_blank(), panel.grid.major.y = element_line(size=.2, color="#656565"), axis.line.x=element_line( size=.3, color="black"),
                       legend.position = "right", legend.key = element_blank(), axis.ticks.y = element_blank(), axis.ticks.x =element_line( size=.3, color="black"),
                       plot.caption=element_text(size=5), axis.text.x=element_text(color="black"))
```

,
                       
                       
                       
                       


```{r historic_size_graph_pic}
ggsave("./HP_pic/historic_size_graph.jpg",width=4, height=3)
```

## Overview of Überhangmandaten for Parties
```{r historic_party_Ueberhang, message=FALSE}
# rename variables for nicer table in readme
 historic_size %>%
          rename(Partei=`Ueberhangmandate_Partei`,
                `Überhangmandate_Anzahl`=`Ueberhangmandate_Anzahl`) %>%
          group_by(Partei) %>%
          filter(!is.na(Partei)) %>%
          summarise(`Überhangmandate (1949-2017)` = sum(Überhangmandate_Anzahl, rm.na=TRUE)) %>%
          kable()
```
## 2017 Election Data Preparation

### get data
```{r 2017_data}
# get data
  raw <- read.csv("https://www.bundeswahlleiter.de/dam/jcr/72f186bb-aa56-47d3-b24c-6a46f5de22d0/btw17_kerg.csv", header=F, sep=";", skip=1, stringsAsFactors = F, na.strings="")
  cleaned <- raw
# keep raw data as backup
  write.csv(raw,"raw.csv", row.names = TRUE)
```

### clean up raw data
```{r 2017_clean, error=FALSE, message=FALSE, warning=FALSE}
# For now there are not really proper column names. So I obtain party names for columns and have all names in row 6 so I can name variables in one line.
# paste party names in front of second vote ("Zweitstimme"); logic: if cell contains "Zweitstimme" grab party name from row 5, 2 cells to the left and paste it in front
# similar logic for 'Erststimme'. 
 
  for(i in 1:ncol(cleaned)){
    if(cleaned[6,i] %in% "Zweitstimmen"){
      cleaned[6,i] <- paste(cleaned[5,i-2], cleaned[6,i])
    }
    if(cleaned[6,i] %in% "Erststimmen"){
      cleaned[6,i] <- paste(cleaned[5,i], cleaned[6,i])
    }
  rm(i)}


# set new column names to names in sixth row
  colnames(cleaned) <- cleaned[6,]
# rename first to third column
  colnames(cleaned)[1:3] <- c("wahlkreisnummer","wahlkreisname", "Bundesland") 
# remove extra rows 
  cleaned <- cleaned[-(1:7),] 
# Drop numbers from previous election (no unnamed); logic: keep all columns whose name contain the letter "n"
  cleaned = cleaned[,grepl("*n",names(cleaned))]
# remove rows with missing values for 'Bundesland' (state) variable as these have NAs for all columns
  cleaned <- cleaned[!(is.na(cleaned$Bundesland)),] 
# every row expect the first two as numeric (we're counting votes, right?)
  cleaned[, 3:97] <- sapply(cleaned[, 3:97], as.numeric)

# drop aggregate values for each state as we care about counties primarily
  cleaned<-cleaned[!(cleaned$`Bundesland`==99),]

#export clean data set as backup
  write.csv(cleaned,"cleaned.csv", row.names = TRUE)
```

### double check all counties included (at the time of writing there are 299 counnties in Germany):
```{r county_check}
nrow(cleaned)
```
## Election Map 

```{r map_actual_election_data}
# Get the highest number of votes for each voting district, as the party with the highest votes gets the mandate. Then rename the parties ( Ionly include the ones included in the Bundestag to save space). Add id for matching with the shape files needed for the map in the next step.
  map_actual_election <- cleaned %>% 
    select(starts_with("wahlkreis") | contains("Erststimme")) %>%
    select(-c(3:6)) %>%
    gather('Partei','Stimmen', `Christlich Demokratische Union Deutschlands Erststimmen`:`Übrige Erststimmen`) %>%
    group_by(wahlkreisnummer) %>%
    top_n(1, Stimmen) %>% 
    mutate(wahlkreisnummer=as.numeric(wahlkreisnummer)) %>%
    mutate(Partei=replace(Partei, Partei=="Christlich Demokratische Union Deutschlands Erststimmen", "CDU"),
          Partei=replace(Partei, Partei=="Sozialdemokratische Partei Deutschlands Erststimmen", "SPD"),
          Partei=replace(Partei, Partei=="DIE LINKE Erststimmen", "LINKE"),
          Partei=replace(Partei, Partei=="BÜNDNIS 90/DIE GRÜNEN Erststimmen", "GRÜNE"),
          Partei=replace(Partei, Partei=="Christlich-Soziale Union in Bayern e.V. Erststimmen", "CSU"),
          Partei=replace(Partei, Partei=="Freie Demokratische Partei Erststimmen", "FDP"),
          Partei=replace(Partei, Partei=="Alternative für Deutschland Erststimmen", "AFD"),
          Partei=replace(Partei, is.numeric(Partei), "Andere")) %>%
    select(starts_with("wahlkreis") | contains("Partei")) %>%
    mutate(id=wahlkreisnummer-1,id=as.character(id))
```

Add shape files
```{r map_actual_election_shape, error=FALSE, message=FALSE, warning=FALSE}
# all election maps borrow heavily from https://interaktiv.morgenpost.de/analyse-bundestagswahl-2017/data/btw17_analysis.html
# load state borders and shapes of the constituencies
  shp_bund <- readOGR("btw17-shapes/bundeslaender_small.shp", "bundeslaender_small", stringsAsFactors=FALSE, encoding="latin1") %>% broom::tidy()
  shp_wahlkreise <- readOGR("btw17-shapes/wahlkreise_small.shp", "wahlkreise_small", stringsAsFactors=FALSE, encoding="latin1") 

  shp_wahlkreise <- shp_wahlkreise %>% broom::tidy() # broom to tidy shape data to make it work smoothly with the graphic package ggplot2
                                                 
  map_actual_election <- merge(shp_wahlkreise, map_actual_election, by="id", all.y=T) # merge shapes with vote data by id
```

## Election Map 2017
```{r map_actual_election, error=FALSE, message=FALSE, warning=FALSE}
# actual graph
  ggplot(data=map_actual_election, aes(x=long, y=lat, group=group))+
    geom_polygon(aes(fill=Partei), show.legend = T) +
    geom_polygon(data=shp_wahlkreise, aes(x=long, y=lat, group=group), fill=NA, color="white", size=0.4) +
    scale_fill_manual(values=c("royalblue1", "#32302e", "blue4", "#46962b", "magenta1", "#E3000F")) +
    labs(title = "Bundestagswahl 2017", subtitle="",caption = "Quelle: Bundeswahlleiter") +
    coord_map() + # apply projection
    theme_void() +  # remove axes
    theme()

```

```{r map_actual_election_pic,echo=FALSE, eval=FALSE}
ggsave("./HP_pic/map_actual_election.jpg",width=4, height=3)
```

# State Data Frames 
Replace state numbers with names create a data set for each state (Land=number to differentiate between states). Also only keep the parties in the parliament

```{r state_seats}
# Vector for the number of seats allocated for each state. The reason for doing this by hand
# stems from the fact that it has already been calculated by the Bundeswahlleiter in the past. Moreover, the formal rule of
# 2 * counties = seats does not always apply due to rounding. 
# source: https://www.bundeswahlleiter.de/dam/jcr/dd81856b-7711-4d9f-98dd-91631ddbc37f/btw17_sitzberechnung.pdf
# The order of seats follows the order of states in the following steps.
  seats <- c(22,12,59,5,128,43,30,76,93,7,24,20,13,32,17,17)
```

### split data frame into data frames for each state. We need this as later on each state is it's own maxmization problem
```{r state_split}
# Additionally replace numbers with names for states for easier understanding. 
# 'state' vector for state names; 'data_state' vector needed for the names of the individual state data frames.
  state <- c("SCH","HAM","NDS","BRE","NRW","HES","RHN","BAD","BAY","SAR","BER","BRA","MEC","SAC","SAA","THU")
  data_state <- paste( "state_",state, sep="")
# create data frames
  for (i in 1:16) {
    cleaned <- cleaned %>% mutate(Bundesland=replace(Bundesland, Bundesland==i, state[i]))
    assign(paste0(data_state[i]), cleaned[cleaned$Bundesland==state[i],])
  rm(i)}
```

### create dynamic state datasets
this is needed for the calculation of the amount of 'Überhangmandaten' and seats per party from a given state

```{r state_dynamic}
dynamic_state <- paste(data_state, "_dynamic", sep="")
for (i in 1:16) {
# create loop dummy datasets so coding gets easier
  loop <-  data.frame(partei =c("CDU","SPD","LINKE","GRÜNE","CSU","FDP","AFD"),
                     stringsAsFactors = FALSE)
# Sitzkontingent state
  loop$Sitzkontingent <- seats[i]
# select state data frame
  loop_state <- get(data_state[i])
# sum Zweitstimmen (second vote)
  loop <- loop %>% mutate(sum_zweitstimmen=sum(loop_state$`Gültige Zweitstimmen`))
# divisor: votes per seat
  loop <- loop %>% mutate(divisor=sum_zweitstimmen/Sitzkontingent)
# Zweitstimmen (second vote) per party
  loop$zweitstimmen <- 0
  loop <- loop %>% mutate(zweitstimmen=replace(zweitstimmen, partei=="CDU", sum(loop_state$`Christlich Demokratische Union Deutschlands Zweitstimmen`, na.rm = TRUE)),
                          zweitstimmen=replace(zweitstimmen, partei=="SPD", sum(loop_state$`Sozialdemokratische Partei Deutschlands Zweitstimmen`, na.rm = TRUE)),         
                          zweitstimmen=replace(zweitstimmen, partei=="LINKE", sum(loop_state$`DIE LINKE Zweitstimmen`, na.rm = TRUE)),                                   
                          zweitstimmen=replace(zweitstimmen, partei=="GRÜNE", sum(loop_state$`BÜNDNIS 90/DIE GRÜNEN Zweitstimmen`, na.rm = TRUE)),                         
                          zweitstimmen=replace(zweitstimmen, partei=="CSU", sum(loop_state$`Christlich-Soziale Union in Bayern e.V. Zweitstimmen`, na.rm = TRUE)),        
                          zweitstimmen=replace(zweitstimmen, partei=="FDP", sum(loop_state$`Freie Demokratische Partei Zweitstimmen`, na.rm = TRUE)),                     
                          zweitstimmen=replace(zweitstimmen, partei=="AFD", sum(loop_state$`Alternative für Deutschland Zweitstimmen`, na.rm = TRUE)))
# save to data frame  
  assign(paste0(dynamic_state[i]), loop,)
# remove not needed values
  rm(loop,loop_state,i)
  }
```

### Get right mandate allocation per party per state for 2017 election. 
Note: for the method described by law you would only add or subtract one (instead of 10). However, this will make the code run significantly slower. If you have time to spare feel free to modify the code.
```{r initial_allocation}
# seats per party
# create vector so you can test later on if seats allocated is how the Bundeswahlleiter determined it to be
  test <- c()
  for (i in 1:16) {
# create loop dummy datasets so coding gets easier
  loop_state <- get(dynamic_state[i])
# votes to mandates per party
  loop_state <- loop_state %>% mutate(mandate=round(zweitstimmen/divisor))
# sum mandates
  number <- sum(loop_state$mandate)
# get mandates allocated to state
  allocated_seats <- seats[i]

# too few allocated seats: solution decrease divisor for convergence from votes to seats
  while (allocated_seats > number){
    loop_state$divisor <- loop_state$divisor-10
    loop_state <- loop_state %>% mutate(mandate=round(zweitstimmen/divisor))
    number <- sum(loop_state$mandate)
  }
# too many allocated seats: solution increase divisor for convergence from votes to seats
  while (number > allocated_seats){
    loop_state$divisor <- loop_state$divisor+10
    loop_state <- loop_state %>% mutate(mandate=round(zweitstimmen/divisor))
    number <- sum(loop_state$mandate)
  }
# add number for checking
  test[i] <- sum(loop_state$mandate)

# save to data frame  
  assign(paste0(dynamic_state[i]), loop_state,)
# clean up  
  rm(loop_state,number,allocated_seats,i)
  
  }
```

### check and clean up
```{r check, error=FALSE, message=FALSE, warning=FALSE}
# check whether amount of seats as calculated by Bundeswahlleiter matches
# with my values per state
  all.equal(seats, test)
# double check: whether mandates by party add up to numbers from Bundeswahlleiter:
# get all mandates per party from each state
  test2 <- state_SCH_dynamic %>% select(partei,mandate)
  for (i in 2:16) {
    loop_state <- get(dynamic_state[i])
    loop_state <- loop_state %>% select(partei,mandate)
    test2 <- rbind(test2,loop_state)
  }
  # get sum and see if they add up to 598
  sum(test2$mandate)
  # get sum of mandates for each party (just cause it's interesting)
  test2 %>% 
    group_by(partei) %>% 
    summarise(mandate = sum(mandate)) %>%
    kable() 

# cleanup
  rm(test,test2)

```

# Optimal Allocation preparation
```{r optimal_allocation}

for (i in 1:16) {

# change loop data set for easy coding  
  loop_state <- get(data_state[i])

# drop all variables for Zweitstimme as we are only concerned with Erststimme from now on
  loop_state <- loop_state[,!grepl("Zweitstimme",names(loop_state))]
  
# only keep parties represented in the parliament as only mandates for these parties will decrease the overall size
loop_state <-  loop_state %>% 
                  select(c(1:3,8:14)) %>% 
                  rename(  CDU = `Christlich Demokratische Union Deutschlands Erststimmen`,
                           SPD = `Sozialdemokratische Partei Deutschlands Erststimmen`,
                           LINKE = `DIE LINKE Erststimmen`,
                           GRÜNE = `BÜNDNIS 90/DIE GRÜNEN Erststimmen`,
                           CSU = `Christlich-Soziale Union in Bayern e.V. Erststimmen`,
                           FDP = `Freie Demokratische Partei Erststimmen`,
                           AFD = `Alternative für Deutschland Erststimmen`)

# get highest amount of Erststimmen for each Wahlkreis: note -1 for parties which had no candidate.
# A negative number is required here as later on when calculating the votes needed, negative numbers
# result in higher numbers, which means they're not selected. (see. next step)
  loop_state[is.na(loop_state)] <- -1

# get hightest number of votes for each Wahlkreis so we can calculate differences   
   loop_state$max <- apply(loop_state[, 4:10], 1, max)

# get difference in votes for each Wahlkreis
        for (x in 1:nrow(loop_state)){
        max <- loop_state[x,11]
        for (y in 4:10){
        loop_state[x,y] <- max-loop_state[x,y]
        }
   }   
  
# sort by wahlkreisnummer - so you can match by rownumber later on
# as matrix from maximizing has no wahlkreisnummer
loop_state <- loop_state %>% 
                  mutate(wahlkreisnummer=as.numeric(wahlkreisnummer)) %>% 
                  arrange(wahlkreisnummer) %>%
                   mutate(matching=1:nrow(loop_state))
   

# save to data frame  
assign(paste0(data_state[i]), loop_state,)

# cleanup
rm(loop_state,i,x,y)    

}

```


# Actual Optimization
```{r optimization}
for (i in 1:16) {
# create loop dummy datasets so coding gets easier
loop_state <- get(dynamic_state[i])

# 1. constrains: Zweitstimmenmandate
  max_seats <- loop_state$mandate 

# keep only variables needed for calculation (difference the from maximum votes)
  loop_state <- get(data_state[i])
  calculation <- loop_state %>% 
                              select(4:10) %>% 
                              as.matrix()


# ----- max sitze pro partei: 1. change order parteien  2. create vector maximale sitze nach zweitstimmen
# ----- https://stackoverflow.com/questions/31813686/lpsolve-in-r-with-character-and-column-sum-contraints  
  
# dimensions matrix
  n_wahlkreise <- nrow(calculation)
  n_parteien <- ncol(calculation)
  ncol = n_wahlkreise*n_parteien 
  
  lp_matching <- make.lp(ncol=ncol)
#we want integer assignments and minimize the votes changed for the optimal solution with no Überhangmandaten
  set.type(lp_matching, columns=1:ncol, type = c("integer"))
  set.objfn(lp_matching, calculation)
  lp.control(lp_matching,sense='max')
 
# X1p + x2p + x3p ... + x28p <= 1 for each wahlkreis
  
# solution vector which contains 1s with length equal to number of wahlkreisen
  max_pro_wahlkreis <- replicate(n_wahlkreise, 1)   
  
  Add_Max_wahlkreis_constraint <- function (wahlkreis_index) {
    wahlkreis_cols <- (0:(n_parteien-1))*n_wahlkreise + wahlkreis_index
    add.constraint(lp_matching, rep(1,n_parteien), indices=wahlkreis_cols,type = c("="), rhs=1)
  }
  
# Add a max_number constraint for each program
  lapply(1:n_wahlkreise, Add_Max_wahlkreis_constraint)
  
# 2. step: max values for each party
  mandate.value <- rep(1, n_wahlkreise) # Add that each mandate is worth on mandate  
  
  Add_max_sitze_constraint <- function (partei_index) {
    partei_cols <- (partei_index-1)*n_wahlkreise + (1:n_wahlkreise) #relevant columns for a given room
    add.constraint(lp_matching, xt=mandate.value, indices=partei_cols, rhs=max_seats[partei_index])
  }
  
# Add a max_number constraint for each party
  lapply(1:n_parteien, Add_max_sitze_constraint)
  
# 10. solve this:
  solve(lp_matching)
  get.variables(lp_matching)  
  
# 11. create dataset (ncol=7  as seven parties in parliament)
  calculation <- get.variables(lp_matching) %>% 
                  matrix(ncol=7)  %>%
                  as.data.frame() %>%
                    rename(CDU_county = 1,
                          SPD_county = 2,
                          LINKE_county = 3,
                          GRÜNE_county = 4,
                          CSU_county = 5,
                          FDP_county = 6,
                          AFD_county = 7)
# 12. variable for matching
  calculation$matching <- 1:nrow(calculation)
  
# 13 .matching
  calculation <- merge(loop_state,calculation,by="matching")  
 
# save to data frame  
assign(paste0(data_state[i]), calculation,)  
  
# clean up
#  rm(i)
}
```




# Creating Variables
```{r optimization_clean_up,warning=FALSE}
for (i in 1:16) {
    clean_up <- get(data_state[i])
# variable indicating winner when optimised
    clean_up <- clean_up %>%
                  mutate(mandate_optimisation="CDU") %>%
                  mutate(mandate_optimisation=replace(mandate_optimisation, SPD_county == 1, "SPD"),
                        mandate_optimisation=replace(mandate_optimisation, LINKE_county == 1, "LINKE"), 
                        mandate_optimisation=replace(mandate_optimisation, GRÜNE_county == 1, "GRÜNE"),
                        mandate_optimisation=replace(mandate_optimisation, CSU_county == 1, "CSU"),
                        mandate_optimisation=replace(mandate_optimisation, FDP_county == 1, "FPD"),
                        mandate_optimisation=replace(mandate_optimisation, AFD_county == 1, "AFD"))
    
    
# get actual direkt Mandate Party name for each Wahlkreis (logic: normal winner = 0)
  
  clean_up <- clean_up %>%
                mutate(mandate_actual="CDU") %>%
                mutate(mandate_actual=replace(mandate_actual, SPD==0, "SPD"),
                      mandate_actual=replace(mandate_actual, LINKE==0, "LINKE"),
                      mandate_actual=replace(mandate_actual, GRÜNE==0, "GRÜNE"),
                      mandate_actual=replace(mandate_actual, CSU==0, "CSU"),
                      mandate_actual=replace(mandate_actual, FDP==0, "FDP"),
                      mandate_actual=replace(mandate_actual, AFD==0, "AFD"))
          
  
# voter change needed for optimisation  
  clean_up <- clean_up %>%
                 mutate(votes_needed=0) %>%
                 mutate(votes_needed=replace(votes_needed, CDU_county == 1, CDU),
                         votes_needed=replace(votes_needed, SPD_county == 1, SPD),
                         votes_needed=replace(votes_needed, LINKE_county == 1, LINKE),
                         votes_needed=replace(votes_needed, GRÜNE_county == 1, GRÜNE),
                         votes_needed=replace(votes_needed, CSU_county == 1, CSU),
                         votes_needed=replace(votes_needed, AFD_county == 1, AFD),
                         votes_needed=replace(votes_needed, FDP_county == 1, FDP)) %>%
                mutate(votes_needed=votes_needed*-1)
  
#  divide by two as half amount would already amount to swing
  clean_up <- clean_up %>% mutate(votes_halved= ceiling(votes_needed/2))
 
}
```
 

# CLEANUP
```{r backup,echo=FALSE, eval=FALSE}
# Zeile 366
# select(-c(13:19))
```
