---
title: "Ueberhangmandate"
output: rmarkdown::github_document
---
  
  ```{r setup, include=FALSE}
knitr::opts_chunk$set(
  fig.path = "README_figs/"
)
```

## Packages

```{r packages, message=FALSE}
# data manipulation
library(tidyr)
library(dplyr)
# optimization
library(lpSolveAPI)
#web scrapping
library(rvest)
#structure replacement Überhang to Ueberhang etc.
library(stringr)
# plots
library(grid)
library(ggplot2)
# packages for maps
library(rgdal)
library(rgeos)
library(mapproj)
# package for table
library(knitr)

```

```{r installs,echo=FALSE, eval=FALSE}
# installs needed in ubuntu 18.04
sudo apt-get install gdal-bin proj-bin libgdal-dev libproj-dev
```


## Theme
```{r theme}
hp_theme <- function(base_size = 13, base_family = "") {
  theme_grey(base_size = base_size, base_family = base_family) %+replace%
    theme(
      
      # Base elements which are not used directly but inherited by others
      line =              element_line(colour = '#DADADA', size = 0.75,
                                       linetype = 1, lineend = "butt"),
      rect =              element_rect(fill = "#F0F0F0", colour = "#F0F0F0",
                                       size = 0.5, linetype = 1),
      text =              element_text(family = base_family, face = "plain",
                                       colour = "#656565", size = base_size,
                                       hjust = 0.5, vjust = 0.5, angle = 0,
                                       lineheight = 0.9, margin = margin(), debug = FALSE),
      plot.margin =       margin(12,10,5,10),
      # Modified inheritance structure of text element
      plot.title =        element_text(size = rel(0.75), family = '' ,
                                       face = 'bold', hjust = 0,
                                       vjust = 2.5, colour = '#3B3B3B'),
      plot.subtitle =     element_text(size = rel(0.4), family = '' ,
                                       face = 'plain', hjust = 0,
                                       vjust = 2.5, colour = '#3B3B3B', margin = margin(0,0,15,0)),
      axis.title.x =      element_blank(),
      axis.title.y =      element_blank(),
      axis.text =         element_text(),
      # Modified inheritance structure of line element
      axis.ticks =        element_line(),
      panel.grid.major =  element_line(),
      panel.grid.minor =  element_blank(),
      
      # Modified inheritance structure of rect element
      plot.background =   element_rect(),
      panel.background =  element_rect(),
      legend.key =        element_rect(colour = '#DADADA'),
      
      # Modifiying legend.position
      legend.position = 'none',
      
      complete = TRUE
    )
}
```

## Historic Size Bundestag Data Set

```{r historic_size_data, error=FALSE, message=FALSE, warning=FALSE}
# read in source
url <- 'https://www.bundeswahlleiter.de/service/glossar/u/ueberhangmandate.html'
# first table [until 2009](before current election reform): no adjustments for Überhangmandate
first_part = url %>%
  read_html() %>%
  html_node(xpath = '/html/body/div/div/main/table[1]') %>%
  html_table(fill = TRUE)
# clean up
# first merge names and first row and clean names up. Second, cut only relevant information
first_part[] <- lapply(first_part, as.character)
names(first_part) <- paste(names(first_part), first_part[1, ], sep = "_")
first_part <- first_part %>%
  rename(`Jahr`=`Jahr der Bundestagswahl_Jahr der Bundestagswahl`,
         `Wahlkreise`=`Zahl der Wahlkreise_Zahl der Wahlkreise`,
         `Mandate`=`Sitze insgesamt 1_Sitze insgesamt 1`) %>%
  slice(2:49) %>%
  select(c(-5)) %>%
  rename_all(funs(str_replace(., "Überhangmandate", "Ueberhangmandate")))

# second table [2013,2017](after current election reform): adjustments for Überhangmandate
second_part = url %>%
  read_html() %>%
  html_node(xpath = '/html/body/div/div/main/table[2]') %>%
  html_table(fill = TRUE)
# clean up
second_part[] <- lapply(second_part, as.character)
names(second_part) <- paste(names(second_part), second_part[1, ], sep = "_")
second_part <- second_part %>%
  rename(`Jahr`=`Jahr der Bundestagswahl_Jahr der Bundestagswahl`,
         `Wahlkreise`=`Zahl der Wahlkreise_Zahl der Wahlkreise`,
         `Mandate`=`Sitze insgesamt_Sitze insgesamt`) %>%
  rename_all(funs(str_replace(., "Überhänge", "Ueberhangmandate"))) %>%
  slice(c(2:8)) %>%
  select(c(1:6))  

# merge data frames  
historic_size <- full_join(first_part,second_part)
historic_size[, 1:5] <- sapply(historic_size[, 1:5], as.numeric)
historic_size <- historic_size%>% mutate(Ueberhangmandate_Partei=replace(Ueberhangmandate_Partei, Ueberhangmandate_Partei=="–", NA)) 
# drop unnecessary data
rm(first_part,second_part, url)
```

## Historic Size Bundestag Graph
```{r historic_size_graph}
historic_size %>%
  distinct(`Jahr`, .keep_all = TRUE) %>%
  select(c(`Jahr`,`Mandate`)) %>%
  ggplot(aes(x=`Jahr`, y=`Mandate`,group=1)) +
  geom_line(aes(group=1), color="#009E73") +
  scale_y_continuous(limits = c(300, 800.1),breaks = c(seq(300,800,100)), expand = c(0, 0), labels=c("300" = "300", "400"="400", "500"="500", "600"= "600", "700"= "700","800"="800 Sitze")) +
  scale_x_continuous(breaks = seq(1949, 2017, by = 4), limits=c(1949,2017), labels=c("1949"="1949", "1953"="53", "1957"="57", "1961"="61", "1965"="65", "1969"="69", "1973"="73", "1977"="77", "1981"="81", "1985"="85", "1989"="89", "1993"="93", "1997"="97", "2001"="01", "2005"="05", "2009"="09", "2013"="13", "2017"="17")) +
  labs(title = "Historische Entwicklung der Größe des Bundestages", subtitle="",caption = "Quelle: Bundeswahlleiter") +
  hp_theme() + theme(axis.text= element_text(size=7.5), axis.title.x = element_blank(),plot.title.position = "plot",  axis.title.y = element_blank(), panel.grid.major.x = element_blank(), panel.grid.major.y = element_line(size=.2, color="#656565"), axis.line.x=element_line( size=.3, color="black"), legend.position = "right", legend.key = element_blank(), axis.ticks.y = element_blank(), axis.ticks.x =element_line( size=.3, color="black"), plot.caption=element_text(size=5), axis.text.x=element_text(color="black"))
```

```{r historic_size_graph_pic}
ggsave("./HP_pic/historic_size_graph.jpg",width=4, height=3)
```

## Overview of Überhangmandaten for Parties
```{r historic_party_Ueberhang, message=FALSE}
 historic_size %>%
    rename(Partei=`Ueberhangmandate_Partei`,
            `Überhangmandate_Anzahl`=`Ueberhangmandate_Anzahl`) %>%
    group_by(Partei) %>%
    filter(!is.na(Partei)) %>%
    summarise(`Überhangmandate (1949-2017)` = sum(Überhangmandate_Anzahl, rm.na=TRUE)) %>%
    kable()
```
## 2017 Election Data Preparation

### get data
```{r 2017_data}
# get data
raw <- read.csv("https://www.bundeswahlleiter.de/dam/jcr/72f186bb-aa56-47d3-b24c-6a46f5de22d0/btw17_kerg.csv", header=F, sep=";", skip=1, stringsAsFactors = F, na.strings="")
# keep raw data
cleaned <- raw
```

### clean up raw data
```{r 2017_clean, error=FALSE, message=FALSE, warning=FALSE}
# paste party names to second vote ("Zweitstimme"); logic if cell contains "Zweitstimme" grab party name from row 5, 2 cells to the left and paste it in front
# similar logic for 'Erststimme'
for(i in 1:ncol(cleaned)){
  if(cleaned[6,i] %in% "Zweitstimmen"){
    cleaned[6,i] <- paste(cleaned[5,i-2], cleaned[6,i])
  }
  if(cleaned[6,i] %in% "Erststimmen"){
    cleaned[6,i] <- paste(cleaned[5,i], cleaned[6,i])
  }
}

# set new column names to names in sixth row
colnames(cleaned) <- cleaned[6,] 
# remove extra rows
cleaned <- cleaned[-(1:7),] 
# rename first to third column
colnames(cleaned)[1:3] <- c("wahlkreisnummer","wahlkreisname", "Bundesland") 
# remove rows with missing values for 'Bundesland' (state) variable
cleaned <- cleaned[!(is.na(cleaned$Bundesland)),] 

# delete every second row from the third row onwards; logic: keep all columns whose name contain the letter "n"
cleaned = cleaned[,grepl("*n",names(cleaned))] 
# every row expect the first two as numeric
number_col <- ncol(cleaned)
for(i in 3:number_col){
  cleaned[[i]] <- as.numeric(cleaned[[i]])
} 

# drop aggregate values for each state
cleaned<-cleaned[!(cleaned$`Bundesland`==99),]

# all columns as numeric
cleaned[, 3:97] <- sapply(cleaned[, 3:97], as.numeric)

# remove not needed values:
rm(i,number_col)

#export clean and raw data set
write.csv(raw,"raw.csv", row.names = TRUE)
write.csv(cleaned,"cleaned.csv", row.names = TRUE)
```

### double check all counties included:
```{r county_check}
nrow(cleaned)
```
## Election Map

Get the highest number of votes for each voting district. Then rename the parties to only include the ones included in the Bundestag to save space. Add id for matching with the shape files.

```{r map_actual_election_data}
map_actual_election <- cleaned %>% 
  select(starts_with("wahlkreis") | contains("Erststimme")) %>%
  select(-c(3:6)) %>%
  gather('Partei','Stimmen', `Christlich Demokratische Union Deutschlands Erststimmen`:`Übrige Erststimmen`) %>%
  group_by(wahlkreisnummer) %>%
  top_n(1, Stimmen) %>% 
  mutate(wahlkreisnummer=as.numeric(wahlkreisnummer)) %>%
  mutate(Partei=replace(Partei, Partei=="Christlich Demokratische Union Deutschlands Erststimmen", "CDU")) %>%
  mutate(Partei=replace(Partei, Partei=="Christlich Demokratische Union Deutschlands Erststimmen", "CDU")) %>%
  mutate(Partei=replace(Partei, Partei=="Sozialdemokratische Partei Deutschlands Erststimmen", "SPD")) %>%
  mutate(Partei=replace(Partei, Partei=="DIE LINKE Erststimmen", "LINKE")) %>%
  mutate(Partei=replace(Partei, Partei=="BÜNDNIS 90/DIE GRÜNEN Erststimmen", "GRÜNE")) %>%
  mutate(Partei=replace(Partei, Partei=="Christlich-Soziale Union in Bayern e.V. Erststimmen", "CSU")) %>%
  mutate(Partei=replace(Partei, Partei=="Freie Demokratische Partei Erststimmen", "FDP")) %>%
  mutate(Partei=replace(Partei, Partei=="Alternative für Deutschland Erststimmen", "AFD")) %>%
  mutate(Partei=replace(Partei, is.numeric(Partei), "Andere")) %>%
  select(starts_with("wahlkreis") | contains("Partei")) %>%
  mutate(id=wahlkreisnummer-1,id=as.character(id))
```

Add shape files
```{r map_actual_election_shape, error=FALSE, message=FALSE, warning=FALSE}
# all election maps borrows heavily from https://interaktiv.morgenpost.de/analyse-bundestagswahl-2017/data/btw17_analysis.html
# load state borders and shapes of the constituencies
shp_bund <- readOGR("btw17-shapes/bundeslaender_small.shp", "bundeslaender_small", stringsAsFactors=FALSE, encoding="latin1") %>% broom::tidy()

wahlkreise <- readOGR("btw17-shapes/wahlkreise_small.shp", "wahlkreise_small", stringsAsFactors=FALSE, encoding="latin1") 

shp_krs <- wahlkreise %>% broom::tidy() # broom to tidy shape data to make it work smoothly with the graphic package ggplot2
rm(wahlkreise)                                                  
map_actual_election <- merge(shp_krs, map_actual_election, by="id", all.y=T) # merge shapes with vote data by id
```

## Election Map 2017
```{r map_actual_election, error=FALSE, message=FALSE, warning=FALSE}
# actual graph
ggplot(data=map_actual_election, aes(x=long, y=lat, group=group))+
  geom_polygon(aes(fill=Partei), show.legend = T) +
  geom_polygon(data=shp_krs, aes(x=long, y=lat, group=group), fill=NA, color="white", size=0.4) +
  scale_fill_manual(values=c("royalblue1", "#32302e", "blue4", "#46962b", "magenta1", "#E3000F")) +
  theme_void() + # remove axes
  coord_map() # apply projection
```

```{r map_actual_election_pic,echo=FALSE, eval=FALSE}
ggsave("./HP_pic/map_actual_election.jpg",width=4, height=3)
```

# State Data Frames 
Replace state numbers with names create a data set for each state (Land=number to differentiate between states). Also only keep the parties in the parliament

```{r state_seats}
# replace numbers with names for states for easier understanding
state <- c("SCH","HAM","NDS","BRE","NRW","HES","RHN","BAD","BAY","SAR","BER","BRA","MEC","SAC","SAA","THU")
data_state <- paste( "state_",state, sep="")
# data frame / vector for the number of seats allocated for each state. The reason for doing this by hand
# stems from the fact that it has already been calculated by the Bundeswahlleiter in the past. Moreover, the formal rule of
# 2 * counties = seats does not always apply due to rounding. 
# source: https://www.bundeswahlleiter.de/dam/jcr/dd81856b-7711-4d9f-98dd-91631ddbc37f/btw17_sitzberechnung.pdf
seats <- c(22,12,59,5,128,43,30,76,93,7,24,20,13,32,17,17)
```

### split data frame into data frames for each state. We need this as later on each state is it's own maxmization problem
```{r state_split}
for (i in 1:16) {
  cleaned <- cleaned %>% mutate(Bundesland=replace(Bundesland, Bundesland==i, state[i]))
  assign(paste0(data_state[i]), cleaned[cleaned$Bundesland==state[i],])
}
# drop values not needed:
rm(i)
```

### create dynamic state datasets
this is needed for the calculation of the amount of Ueberhangmandaten and seats per party from a given state

```{r state_dynamic}
dynamic_state <- paste( "dynamic_",state, sep="")

for (i in 1:16) {
  # create loop dummy datasets so coding gets easier
  loop <-  data.frame(partei =c("CDU","SPD","LINKE","GRÜNE","CSU","FDP","AFD"),
                     stringsAsFactors = FALSE)
  loop_state <- get(data_state[i])
  # Sitzkontingent state
  loop$Sitzkontingent <- seats[i]
  # sum Zweitstimmen
  loop <- loop %>% mutate(sum_zweitstimmen=sum(loop_state$`Gültige Zweitstimmen`))
  # divisor: votes per seat
  loop <- loop %>% mutate(divisor=sum_zweitstimmen/Sitzkontingent)
  # Zweitstimmen per party
  loop$zweitstimmen <- 0
  loop <- loop %>% mutate(zweitstimmen=replace(zweitstimmen, partei=="CDU", sum(loop_state$`Christlich Demokratische Union Deutschlands Zweitstimmen`, na.rm = TRUE))) 
  loop <- loop %>% mutate(zweitstimmen=replace(zweitstimmen, partei=="SPD", sum(loop_state$`Sozialdemokratische Partei Deutschlands Zweitstimmen`, na.rm = TRUE)  ))         
  loop <- loop %>% mutate(zweitstimmen=replace(zweitstimmen, partei=="LINKE", sum(loop_state$`DIE LINKE Zweitstimmen`, na.rm = TRUE)  ))                                     
  loop <- loop %>% mutate(zweitstimmen=replace(zweitstimmen, partei=="GRÜNE", sum(loop_state$`BÜNDNIS 90/DIE GRÜNEN Zweitstimmen`, na.rm = TRUE)  ))                         
  loop <- loop %>% mutate(zweitstimmen=replace(zweitstimmen, partei=="CSU", sum(loop_state$`Christlich-Soziale Union in Bayern e.V. Zweitstimmen`, na.rm = TRUE)  ))         
  loop <- loop %>% mutate(zweitstimmen=replace(zweitstimmen, partei=="FDP", sum(loop_state$`Freie Demokratische Partei Zweitstimmen`, na.rm = TRUE)  ))                      
  loop <- loop %>% mutate(zweitstimmen=replace(zweitstimmen, partei=="AFD", sum(loop_state$`Alternative für Deutschland Zweitstimmen`, na.rm = TRUE)  ))
  # save to data frame  
  assign(paste0(dynamic_state[i]), loop,)
  # remove not needed values
  rm(loop,loop_state,i)
  }
```

Get right mandate allocation per party per state for 2017 election. Note: in the method described by law you would only add or subtract one (instead of 10). However, this will make the code run significantly longer. If you have time to spare feel free to modify the code.
```{r initial_allocation}
  # 6. Sitze der jeweiligen Parteien:
# create vector so you can test later on if seats allocated is how the Bundeswahlleiter determined it to be
test <- c()
for (i in 1:16) {
# create loop dummy datasets so coding gets easier
loop_state <- get(dynamic_state[i])
# votes to mandates per party
loop_state <- loop_state %>% mutate(mandate=round(zweitstimmen/divisor))
# sum mandates
number <- sum(loop_state$mandate)
# get mandates allocated to state
allocated_seats <- seats[i]

  # too few allocated seats: solution decrease divisor for convergence from votes to seats
  while (allocated_seats > number){
    loop_state$divisor <- loop_state$divisor-10
    loop_state <- loop_state %>% mutate(mandate=round(zweitstimmen/divisor))
    number <- sum(loop_state$mandate)
  }
  # too many allocated seats: solution increase divisor for convergence from votes to seats
  while (number > allocated_seats){
    loop_state$divisor <- loop_state$divisor+10
    loop_state <- loop_state %>% mutate(mandate=round(zweitstimmen/divisor))
    number <- sum(loop_state$mandate)
  }
# add number for checking
test[i] <- sum(loop_state$mandate)

# save to data frame  
assign(paste0(dynamic_state[i]), loop_state,)

}
```
check and clean up
```{r check}
# check whether operation successful
all.equal(seats, test)

# cleanup
rm(loop_state,number,allocated_seats,test,i)

```

# Optimal Allocation preparation
```{r optimal_allocation}

for (i in 1:16) {

# change loop data set for easy coding  
  loop_state <- get(data_state[i])

# drop all variables for Zweitstimme as we are only concerned with Erststimme from now on
  loop_state <- loop_state[,!grepl("Zweitstimme",names(loop_state))]
  
# only keep parties represented in the parliament as only mandates for these parties will decrease the overall size
loop_state <-  loop_state %>% 
                  select(c(1:3,8:14)) %>% 
                  rename(  CDU = `Christlich Demokratische Union Deutschlands Erststimmen`,
                           SPD = `Sozialdemokratische Partei Deutschlands Erststimmen`,
                           LINKE = `DIE LINKE Erststimmen`,
                           GRÜNE = `BÜNDNIS 90/DIE GRÜNEN Erststimmen`,
                           CSU = `Christlich-Soziale Union in Bayern e.V. Erststimmen`,
                           FDP = `Freie Demokratische Partei Erststimmen`,
                           AFD = `Alternative für Deutschland Erststimmen`)

# get highest amount of Erststimmen for each Wahlkreis: note -1 for parties which had no candidate.
# A negative number is required here as later on when calculating the votes needed, negative numbers
# result in higher numbers, which means they're not selected. (see. next step)
  loop_state[is.na(loop_state)] <- -1

# get hightest number of votes for each Wahlkreis so we can calculate differences   
   loop_state$max <- apply(loop_state[, 4:10], 1, max)

# get difference in votes for each Wahlkreis
        for (x in 1:nrow(loop_state)){
        max <- loop_state[x,11]
        for (y in 4:10){
        loop_state[x,y] <- loop_state[x,y]-max
        }
   }   
  
# sort by wahlkreisnummer - so you can match by rownumber later on
# as matrix from maximizing has no wahlkreisnummer
loop_state <- loop_state %>% 
                  mutate(wahlkreisnummer=as.numeric(wahlkreisnummer)) %>% 
                  arrange(wahlkreisnummer) %>%
                   mutate(matching=1:nrow(loop_state))
   

# save to data frame  
assign(paste0(data_state[i]), loop_state,)

# cleanup
rm(loop_state,i,x,y)    

}

```


# Actual Optimization
```{r optimization}
for (i in 1:16) {
# create loop dummy datasets so coding gets easier
loop_state <- get(dynamic_state[i])

# 1. constrains: Zweitstimmenmandate
  max_seats <- loop_state$mandate 

# keep only variables needed for calculation (difference the from maximum votes)
  loop_state <- get(data_state[i])
  calculation <- loop_state %>% 
                              select(4:10) %>% 
                              as.matrix()


# ----- max sitze pro partei: 1. change order parteien  2. create vector maximale sitze nach zweitstimmen
# ----- https://stackoverflow.com/questions/31813686/lpsolve-in-r-with-character-and-column-sum-contraints  
  
# dimensions matrix
  n_wahlkreise <- nrow(calculation)
  n_parteien <- ncol(calculation)
  ncol = n_wahlkreise*n_parteien 
  
  lp_matching <- make.lp(ncol=ncol)
#we want integer assignments and minimize the votes changed for the optimal solution with no Überhangmandaten
  set.type(lp_matching, columns=1:ncol, type = c("integer"))
  set.objfn(lp_matching, calculation)
  lp.control(lp_matching,sense='max')
 
# X1p + x2p + x3p ... + x28p <= 1 for each wahlkreis
  
# solution vector which contains 1s with length equal to number of wahlkreisen
  max_pro_wahlkreis <- replicate(n_wahlkreise, 1)   
  
  Add_Max_wahlkreis_constraint <- function (wahlkreis_index) {
    wahlkreis_cols <- (0:(n_parteien-1))*n_wahlkreise + wahlkreis_index
    add.constraint(lp_matching, rep(1,n_parteien), indices=wahlkreis_cols,type = c("="), rhs=1)
  }
  
# Add a max_number constraint for each program
  lapply(1:n_wahlkreise, Add_Max_wahlkreis_constraint)
  
# 2. step: max values for each party
  mandate.value <- rep(1, n_wahlkreise) # Add that each mandate is worth on mandate  
  
  Add_max_sitze_constraint <- function (partei_index) {
    partei_cols <- (partei_index-1)*n_wahlkreise + (1:n_wahlkreise) #relevant columns for a given room
    add.constraint(lp_matching, xt=mandate.value, indices=partei_cols, rhs=max_seats[partei_index])
  }
  
# Add a max_number constraint for each party
  lapply(1:n_parteien, Add_max_sitze_constraint)
  
# 10. solve this:
  solve(lp_matching)
  get.variables(lp_matching)  
  
# 11. create dataset (ncol=7  as seven parties in parliament)
  calculation <- get.variables(lp_matching) %>% 
                  matrix(ncol=7)  %>%
                  as.data.frame() %>%
                    rename(CDU_county = 1,
                          SPD_county = 2,
                          LINKE_county = 3,
                          GRÜNE_county = 4,
                          CSU_county = 5,
                          FDP_county = 6,
                          AFD_county = 7)
# 12. variable for matching
  calculation$matching <- 1:nrow(calculation)
  
# 13 .matching
  calculation <- merge(loop_state,calculation,by="matching")  
 
# save to data frame  
assign(paste0(data_state[i]), calculation,)  
  
# clean up
#  rm(i)
}
```




# Creating Variables
```{r optimization_clean_up,warning=FALSE}
for (i in 1:16) {
    clean_up <- get(data_state[i])
# variable indicating winner when optimised
    clean_up <- clean_up %>%
                  mutate(mandate_optimisation="CDU") %>%
                  mutate(mandate_optimisation=replace(mandate_optimisation, SPD_county == 1, "SPD"),
                        mandate_optimisation=replace(mandate_optimisation, LINKE_county == 1, "LINKE"), 
                        mandate_optimisation=replace(mandate_optimisation, GRÜNE_county == 1, "GRÜNE"),
                        mandate_optimisation=replace(mandate_optimisation, CSU_county == 1, "CSU"),
                        mandate_optimisation=replace(mandate_optimisation, FDP_county == 1, "FPD"),
                        mandate_optimisation=replace(mandate_optimisation, AFD_county == 1, "AFD"))
    
    
# get actual direkt Mandate Party name for each Wahlkreis (logic: normal winner = 0)
  
  clean_up <- clean_up %>%
                mutate(mandate_actual="CDU") %>%
                mutate(mandate_actual=replace(mandate_actual, SPD==0, "SPD"),
                      mandate_actual=replace(mandate_actual, LINKE==0, "LINKE"),
                      mandate_actual=replace(mandate_actual, GRÜNE==0, "GRÜNE"),
                      mandate_actual=replace(mandate_actual, CSU==0, "CSU"),
                      mandate_actual=replace(mandate_actual, FDP==0, "FDP"),
                      mandate_actual=replace(mandate_actual, AFD==0, "AFD"))
          
  
# voter change needed for optimisation  
  clean_up <- clean_up %>%
                 mutate(votes_needed=0) %>%
                 mutate(votes_needed=replace(votes_needed, CDU_county == 1, CDU),
                         votes_needed=replace(votes_needed, SPD_county == 1, SPD),
                         votes_needed=replace(votes_needed, LINKE_county == 1, LINKE),
                         votes_needed=replace(votes_needed, GRÜNE_county == 1, GRÜNE),
                         votes_needed=replace(votes_needed, CSU_county == 1, CSU),
                         votes_needed=replace(votes_needed, AFD_county == 1, AFD),
                         votes_needed=replace(votes_needed, FDP_county == 1, FDP)) %>%
                mutate(votes_needed=votes_needed*-1)
  
#  divide by two as half amount would already amount to swing
  clean_up <- clean_up %>% mutate(votes_halved= ceiling(votes_needed/2))
 
}
```
 

# CLEANUP
```{r backup,echo=FALSE, eval=FALSE}
# Zeile 366
# select(-c(13:19))
```
